Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-22 paper.tex

The compression was performed on the All of Us Researcher Workbench platform 
(based on Google Cloud), using a single compute node with 64 CPUs and 416 GiB 
of RAM, which cost \$3.80/hour. We first converted the VCF to an ICF, which 
took 6 hours and 17 minutes of effective runtime and utilized 343 hours of CPU 
time, achieving a parallel efficiency of 86\%. This step incurred a cost of 
\$23.88. The resulting ICF representation required 7.5 GiB of storage distributed 
across 91,232 files. Next, the ICF was converted to Zarr, 
after removing the field \texttt{call\_FT}, which is 
suspected to be problematic. The conversion took 53 minutes and 40 seconds, with 
a CPU time of 52 hours and 34 minutes and a parallel efficiency of 92\%, at a 
cost of \$3.40. This second Zarr representation occupied 8 GiB of storage across 
214,190 chunk files.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns produced 
    from the dataset using \texttt{vcf2zarr} with chunk sizes of 1,000 variants and 
    10,000 samples. For column details, see Table~\ref{tab-genomics-england-data}.
    \label{tab-top-fields-data}}
\begin{tabular}{llS[table-format=3.2]S[table-format=4.1]S[table-format=3.2]}
\toprule
{Field} & {Type} & {Storage} & {Compress} & {\% Total} \\
\midrule
/call_LAD & int16 & 2.26 GiB & 290.0 & 29.63% \\
/call_GQ & int8 & 1.89 GiB & 87.0 & 24.78% \\
/call_LA & int8 & 1.24 GiB & 260.0 & 16.26% \\
/call_genotype & int8 & 914.22 MiB & 370.0 & 11.70% \\
/call_RGQ & int16 & 729.57 MiB & 460.0 & 9.34% \\
/call_genotype_mask & bool & 606.02 MiB & 550.0 & 7.76% \\
\bottomrule
\end{tabular}
\end{table}

>>>>

The compression was performed on the All of Us Researcher Workbench platform 
(based on Google Cloud), using a single compute node with 64 CPUs and 416 GiB 
of RAM, which cost \$3.80/hour. We first converted the VCF to an ICF, which 
took 6 hours and 17 minutes of effective runtime and utilized 343 hours of CPU 
time, achieving a parallel efficiency of 86\%. This step incurred a cost of 
\$23.88. The resulting ICF representation required 7.5 GiB of storage distributed 
across 91,232 files. Next, the ICF was converted to Zarr, which took 1 hour and 8 minutes, with 
a CPU time of 68 hours and 8 minutes and a parallel efficiency of 94\%, at a 
cost of \$4.31. This second Zarr representation occupied 8.5 GiB of storage across 
232,809 chunk files.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns produced 
    from the dataset using \texttt{vcf2zarr} with chunk sizes of 1,000 variants and 
    10,000 samples. For column details, see Table~\ref{tab-genomics-england-data}.
    \label{tab-top-fields-data}}
\begin{tabular}{llS[table-format=3.2]S[table-format=4.1]S[table-format=3.2]}
\toprule
{Field} & {Type} & {Storage} & {Compress} & {\% Total} \\
\midrule
/call_LAD & int16 & 2.26 GiB & 290.0 & 27.85% \\
/call_GQ & int8 & 1.89 GiB & 87.0 & 23.29% \\
/call_LA & int8 & 1.24 GiB & 260.0 & 15.28% \\
/call_genotype & int8 & 914.22 MiB & 370.0 & 11.00% \\
/call_RGQ & int16 & 729.57 MiB & 460.0 & 8.78% \\
/call_genotype_mask & bool & 606.02 MiB & 550.0 & 7.29% \\
/call_FT & object & 498.61 MiB & 2700.0 & 6.00% \\
\bottomrule
\end{tabular}
\end{table}
